---
title: "Final Project- Data Management"
author: "Sierrah Mueller"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(DHARMa)
library(MASS)
library(pscl)
library(AER)
library(tidyverse)
library(readxl)
```

# Chapter 2 RMD: Juvenile Redtail Parrotfish Habitat 

## Research Question 1
### Is there a difference in Redtail Parrotfish Density Between Habitat Types and Size Classes

```{r, include=FALSE, echo = FALSE}

abundance_dfnewimport <- read_excel("Data/CombinedData_Abundance-Habitat.xlsx", sheet = "OverallHabitatANOVA")
abundance_df <- abundance_dfnewimport
abundance_df %>%
  mutate(totalabun = rowSums(.[2:9]))

```


## Research Question 2

### How does benthic substrate percent cover affect the density of Redtail Parrotfish? 
Negative Binomial Regression
```{r NBR: Substrate, echo= FALSE, message = FALSE}
glm_df <- read_excel("Data/glm_df.xlsx")

poisson_benthic <- glm(Abundance ~ MacaUNSP + Dictyota + TURF + SEAGRASS + CORAL + SPONGE, family = "poisson", data = glm_df)
summary(poisson_benthic)
dispersiontest(poisson_benthic)
m_benthic <- glm.nb(Abundance ~ MacaUNSP + Dictyota + TURF + SEAGRASS + CORAL + SPONGE , data = glm_df)
summary(m_benthic)
odTest(m_benthic)


```

### How do complexity and depth affect the density of Redtail Parrotfish?
```{r NBR, echo= FALSE, message = FALSE}
nbr_complexdepth <- read_excel("Data/nbr_complexdepth.xlsx")
noout <- subset(nbr_complexdepth,Complexity < 100)



```

