---
title: "Final Project- Data Management"
author: "Sierrah Mueller"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Chapter 2 RMD: Juvenile Redtail Parrotfish Habitat 

## Research Question 1:
### Is there a difference in Redtail Parrotfish Density Between Habitat Types and Size Classes

```{r, include=FALSE}
library(tidyverse)
library(readxl)
abundance_dfnewimport <- read_excel("Data/CombinedData_Abundance-Habitat.xlsx", sheet = "OverallHabitatANOVA")
abundance_df <- abundance_dfnewimport
abundance_df %>%
  mutate(totalabun = rowSums(.[2:9]))

```


#Research Question 2: 
### How does benthic substrate percent cover affect the density of Redtail Parrotfish? 
Negative Binomial Regression
```{r NBR: Substrate}
library(readxl)
library(DHARMa)
library(MASS)
library(pscl)
library(AER)
glm_df <- read_excel("Data/glm_df.xlsx")
poisson_benthic <- glm(Abundance ~ MacaUNSP + Dictyota + TURF + SEAGRASS + CORAL + SPONGE, family = "poisson", data = glm_df)
summary(poisson_benthic)
dispersiontest(poisson_benthic)


