---
title: "Thesis RMD"
author: "Sierrah Mueller"
date: "2023-10-26"
output: html_document
---
#would like to make each research questions a new tab with headers explaining analysis in each.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 2: Juvenile Redtail Parrotfish Habitat 

## Research quesetion 1:
### Is there a difference in Redtail Parrotfish Density Between Habitat Types and Size Classes

Kruskall-Wallis Test
```{r Abundance~Habitat}
library(tidyverse)
library(readxl)
abundance_dfnewimport <- read_excel("Data/CombinedData_Abundance-Habitat.xlsx", sheet = "OverallHabitatANOVA")
abundance_df <- abundance_dfnewimport %>%
  mutate(totalabun = rowSums(.[2:9]))
density_df <- abundance_df %>%
  mutate(density = totalabun*2) %>%
  mutate(MajorHabitat = ifelse(MajorHabitat == "Seagrass", "Seagrass Fringe", MajorHabitat))
# density_df
density_df
kruskal.test(density ~ MajorHabitat, data = density_df)
pairwise <- pairwise.wilcox.test(density_df$density, density_df$MajorHabitat,
                     p.adjust.method = "BH")
pairwise

fig1 <- ggplot(density_df, aes(x = MajorHabitat, y = density)) +
  geom_boxplot(outlier.colour = "grey", outlier.shape = 16, outlier.size = 1) +
  geom_jitter(width = 0.3, size = 1, alpha = 0.2) +
  labs(y=bquote("Density (N per 100 m"^"2"*")")) +
  xlab("Habitat Type") +
  theme(axis.text.x=element_text(size=9))
fig1
```


Chi-Square Test
```{r Chi-Square Test}
#0-3
habitat <- c("Hardbottom", "Rocky Reef", "Seagrass")
size_class <- c("0-3")
count <- c(74, 70, 54)
contingency_table <- matrix(count, nrow = length(habitat), ncol = length(size_class))
rownames(contingency_table) <- habitat
colnames(contingency_table) <- size_class
print(contingency_table)
size0_3 <- chisq.test(contingency_table)
size0_3
#Not significant: 0.183

#4-5
habitat <- c("Hardbottom", "Rocky Reef", "Seagrass")
size_class <- c("4-5")
count <- c(275, 187 , 118)
contingency_table <- matrix(count, nrow = length(habitat), ncol = length(size_class))
rownames(contingency_table) <- habitat
colnames(contingency_table) <- size_class
size4_5 <- chisq.test(contingency_table)
size4_5
#Significant: < 0.001

#6-10
habitat <- c("Hardbottom", "Rocky Reef", "Seagrass")
size_class <- c("4-5")
count <- c(303, 245, 125)
contingency_table <- matrix(count, nrow = length(habitat), ncol = length(size_class))
rownames(contingency_table) <- habitat
colnames(contingency_table) <- size_class
size6_10 <- chisq.test(contingency_table)
size6_10
#Significant: < 0.001

#11-15
habitat <- c("Hardbottom", "Rocky Reef", "Seagrass")
size_class <- c("11-15")
count <- c(88, 86, 23)
contingency_table <- matrix(count, nrow = length(habitat), ncol = length(size_class))
rownames(contingency_table) <- habitat
colnames(contingency_table) <- size_class
size11_15 <- chisq.test(contingency_table)
size11_15
#Significant: < 0.001

#16-20
habitat <- c("Hardbottom", "Rocky Reef", "Seagrass")
size_class <- c("16-20")
count <- c(28, 47, 17)
contingency_table <- matrix(count, nrow = length(habitat), ncol = length(size_class))
rownames(contingency_table) <- habitat
colnames(contingency_table) <- size_class
size16_20 <- chisq.test(contingency_table)
size16_20
#Signficiant: < 0.001

#21-30
habitat <- c("Hardbottom", "Rocky Reef", "Seagrass")
size_class <- c("16-20")
count <- c(7, 12, 2)
contingency_table <- matrix(count, nrow = length(habitat), ncol = length(size_class))
rownames(contingency_table) <- habitat
colnames(contingency_table) <- size_class
size21_30<- chisq.test(contingency_table)
size21_30
#Significant: < 0.001
```

###Preliminary Roving Fish Survey Graph
```{r Prelim Fish Suvey Graph}
#Prelim Analysis
melt_prelim <- melt(prelim_overall, id = c("Habitat"))

prelim_fig <- ggplot(melt_prelim, aes(x=Habitat, y= value, fill = variable))  +
  geom_col(position = position_dodge(),
           width = (0.75)) +
  labs(y=bquote("Abundance"),
       fill = "Size Class (cm)") +
  xlab("Habitat Type") +
  scale_fill_brewer(palette = "Dark2")

prelim_fig
```


#Research Question 2: 
### How does benthic substrate percent cover affect the density of Redtail Parrotfish? 
Negative Binomial Regression
```{r NBR: Substrate}
library(readxl)
glm_df <- read_excel("Data/glm_df.xlsx")
poisson_benthic <- glm(Abundance ~ MacaUNSP + Dictyota + TURF + SEAGRASS + CORAL + SPONGE, family = "poisson", data = glm_df)
summary(poisson_benthic)
dispersiontest(poisson_benthic)
library(DHARMa)
resp <- simulateResiduals(modelpois, refit = T)
testDispersion(resp, plot = F)
plot(resp)
#Plot shows significant dispersion (3.43, p= <0.001)

m_benthic <- glm.nb(Abundance ~ MacaUNSP + Dictyota + TURF + SEAGRASS + CORAL + SPONGE , data = glm_df)
summary(m_benthic)
m2_benthic <- update(m_benthic, . ~ . - Dictyota)
anova(m_benthic, m2_benthic)
pchisq(2 * (logLik(m1) - logLik(m3)), df = 1, lower.tail = FALSE)
qchisq(p=5.866e-69, df=8, lower.tail=FALSE)
#Chi-square 341.5115 which means NBR model is better

library(pscl)
odTest(m_benthic)
#Chi-square Test Statistic = 308.092, p < 0.0001
library(AER)
dispersiontest(poisson_benthic)
#Dispersion 3.352

#Reference website: https://www.uio.no/studier/emner/matnat/math/STK3100/h23/undervisningsmateriale/exercise_solutions/exercise_7.31.pdf
#Another: https://bookdown.org/ks6017/GLM_bookdown3/chapter-4-poisson-regression-and-extensions.html#poisson-distribution


plot_model(m_benthic, type = "pred", terms = "Dictyota", axis.title = "Density", title = "")

#get confidence interval for complexity
CI_ben <- ggpredict(m_benthic, interval = "confidence", type = "fixed",
                 terms = "Dictyota")
CI_ben <- CI_ben %>% as.data.frame()

ggplot() +
  geom_point(data = glm_df, aes(x = Dictyota,
                               y = Abundance,
                              ),
             alpha = 0.2) +
  geom_path(data = CI_ben, aes(x = x, y = predicted)) +
  geom_ribbon(data = CI_ben, aes(x = x, ymin = conf.low,
                                  ymax = conf.high),
              alpha = 0.5)

```

### How does benthic substrate percent cover affect the density of Redtail Parrotfish within each habitat type? 
Negative Binomial Regression
```{r NBR: Substrate x Habitat Type}
#SEAGRASS FRINGE

df_sg = glm_df %>%
  filter(Habitat == "Seagrass")
m_poisson_sg <- glm(Abundance ~  MacaUNSP + TURF + Dictyota + THALASSIA + SYRINGODIUM + HALOPHILA, family = "poisson", data = df_sg)
m_sg <- glm.nb(Abundance ~  MacaUNSP + TURF + Dictyota + THALASSIA + SYRINGODIUM + HALOPHILA, data = df_sg)
summary(m_sg)
summary(m_poisson_sg)
dispersiontest(m_poisson_sg)
odTest(m_sg)

#ROCKY REEF

df_rr = glm_df %>%
  filter(Habitat == "Rocky Reef")
m_rr <- glm.nb(Abundance ~  MacaUNSP + Dictyota + TURF + CORAL + SPONGE, data = df_rr)
m_poisson_rr <- glm(Abundance ~  MacaUNSP + Dictyota + TURF + CORAL + SPONGE, family = "poisson", data = df_rr)
dispersiontest(m_poisson_rr)
odTest(m_rr)
summary(m_rr)

#HARDBOTTOM
df_hb = glm_df %>%
  filter(Habitat == "Hardbottom")
m_hb <- glm.nb(Abundance ~  MacaUNSP + Dictyota + TURF + CORAL + SPONGE, data = df_hb)
m_poisson_hb <- glm(Abundance ~  MacaUNSP + Dictyota + TURF + CORAL + SPONGE, family = "poisson", data = df_hb)
dispersiontest(m_poisson_hb)
odTest(m_hb)
summary(m_hb)


```

### How do complexity and depth affect the density of Redtail Parrotfish?
```{r NBR: ComplexityxDepth}
noout <- subset(nbr_complexdepth,Complexity < 100)
modelpois <- glm(Density ~ Complexity + AvgDepth, family = "poisson", data = noout)
summary(modelpois)
library(DHARMa)
resp <- simulateResiduals(modelpois, refit = T)
testDispersion(resp, plot = F)
plot(resp)
#SHOWS DISPERSION

m_depthcomplex <- glm.nb(Density ~ Complexity + AvgDepth, data = noout)
summary(m_depthcomplex)
m2_depthcomplex <- update(m_depthcomplex, . ~ . - Complexity)
anova(m_depthcomplex, m2_depthcomplex)
poisson_depthcomplex <- glm(Density ~ Complexity + AvgDepth, family = "poisson", data = noout)
install.packages("AER")
library(AER)
dispersiontest(poisson_depthcomplex)
#Dispersion 3.441745 
pchisq(2 * (logLik(m1) - logLik(m3)), df = 1, lower.tail = FALSE)
qchisq(p=5.16e-84, df=4, lower.tail=FALSE)
#Chi-square test 394.2 - ABOVE CRITICAL VALUE MEANS GOOD TEST
qchisq(0.05, df=4, lower.tail=FALSE)

library(pscl)
odTest(m_depthcomplex)
#Critical value of test statistic at the alpha= 0.05 level: 2.7055 
#Chi-Square Test Statistic =  347.7721 p-value = < 2.2e-16 

plot_model(m_depthcomplex, type = "pred", terms = "Complexity", axis.title = "Density Prediction", title = "")

#get confidence interval for complexity
CI_comp <- ggpredict(m_depthcomplex, interval = "confidence", type = "fixed",
                 terms = "Complexity")
CI_comp <- CI_comp %>% as.data.frame()

ggplot() +
  geom_point(data = noout, aes(x = Complexity,
                               y = Density,
                              ),
             alpha = 0.2) +
  geom_path(data = CI_comp, aes(x = x, y = predicted)) +
  geom_ribbon(data = CI_comp, aes(x = x, ymin = conf.low,
                                  ymax = conf.high),
              alpha = 0.5)




```

### How do complexity affect the density of Redtail Parrotfish within each size class?

```{r NBR: Complexity ~ Size Class}
m1_poisson <- glm(size1 ~ Complexity, family = "poisson", data = noout)
m1 <- glm.nb(size1 ~ Complexity , data = noout)
s1 <- summary(m1)
s1
plot_model(m1, type = "pred", terms = "Complexity", axis.title = "Density Prediction", title = "", show.data = TRUE)
predicted_values <- predict(m1, type = "response", se.fit = TRUE)
data_with_predictions <- cbind(noout, predicted_values)

#0-3 not significant
odTest(m1)
dispersiontest(m1_poisson)

m2 <- glm.nb(size2 ~ Complexity , data = noout)
m2_poisson <- glm(size2 ~ Complexity, family = "poisson", data = noout)
s2 <- summary(m2)
s2
plot_model(m2, type = "pred", terms = "complexity", axis.title = "Density Prediction", title = "")
#4-5 not significant
odTest(m2)
dispersiontest(m2_poisson)

m3_poisson <- glm(size3 ~ Complexity, family = "poisson", data = noout)
m3 <- glm.nb(size3 ~ Complexity , data = noout)
s3 <- summary(m3)
s3
plot_model(m3, type = "pred", terms = "complexity", axis.title = "Density Prediction", title = "")
#6-10 significant
odTest(m3)
dispersiontest(m3_poisson)


m4_poisson <- glm(size4 ~ Complexity, family = "poisson", data = noout)
m4 <- glm.nb(size4 ~ Complexity , data = noout)
s4 <- summary(m4)
s4
plot_model(m4, type = "pred", terms = "complexity", axis.title = "Density Prediction", title = "")
#11-15 significant
odTest(m4)
dispersiontest(m4_poisson)

m_5poisson <- glm(size5 ~ Complexity, family = "poisson", data = noout)
m5 <- glm.nb(size5 ~ Complexity , data = noout)
s5 <- summary(m5)
s5
plot_model(m6, type = "pred", terms = "complexity", axis.title = "Density Prediction", title = "")
#16-20 significant
odTest(m5)
dispersiontest(m5_poisson)

m6_poisson <- glm(size6 ~ Complexity, family = "poisson", data = noout)
summary(m6_poisson)
m6 <- glm.nb(size6 ~ Complexity , data = noout)
s6 <- summary(m6)
s6
plot_model(m6, type = "pred", terms = "complexity", axis.title = "Density Prediction", title = "")
#21-30 significant but bad 
odTest(m6)
dispersiontest(m6_poisson)

m7_poisson <- glm(size7 ~ Complexity, family = "poisson", data = noout)
summary(m7_poisson)
m7 <- glm.nb(size7 ~ Complexity , data = noout)
s7 <- summary(m7)
s7
plot_model(m6, type = "pred", terms = "complexity", axis.title = "Density Prediction", title = "")
#21-30 significant but bad 
odTest(m7)
dispersiontest(m7_poisson)

```

###Is there a difference in complexity between habitat types?
```{r Complexity ~ Habitat type}
complex_sum  <- ddply(nbr_complexdepth, ~MajorHabitat, summarise,
                      mean = mean(Complexity), 
                      sd = sd(Complexity), 
                      n = length(Complexity), 
                      SEM = sd(Complexity)/sqrt
                      (length(Complexity)))

kruskal.test(Complexity ~ MajorHabitat, data = nbr_complexdepth)
#Kruskal-Wallis chi-squared = 129.35, df = 2, p-value <0.001 
#Significant Different Between Complexity in the Major Habitat Types
dunnTest(Complexity ~ MajorHabitat,
        data=nbr_complexdepth,
        method="bonferroni")

pairwise.wilcox.test(nbr_complexdepth$Complexity, nbr_complexdepth$MajorHabitat, 
                     p.adjust.method = "BH")

sigs <- c("a", "b", "c")

fig_complex <- ggplot(nbr_complexdepth, aes(x = MajorHabitat, y = Complexity)) +
  geom_boxplot(outlier.colour = "grey", outlier.shape = 16, outlier.size = 1) +
  geom_jitter(width = 0.3, size = 1, alpha = 0.2) +
  labs(y=bquote("Topographic Complexity (cm)")) +
  xlab("Habitat Type") +
  theme(axis.text.x=element_text(size=9))
fig_complex

nbr_complexdepth$MajorHabitat[nbr_complexdepth$MajorHabitat == "Seagrass"] <- "Seagrass Fringe"
```


###Benthic Biological Substrate Proportion
```{r Percent Cover Substrate}
benthic <- benthicdf %>%
  subset(select = c(Habitat, CORAL, SPONGE, TURF, Nonliving, SEAGRASS, allother, MacaUNSP, DICT)) %>%
  rename("Coral" = "CORAL",
         "Sponge" = "SPONGE", 
         "Turf Algae" = "TURF", 
         "Non-living" = "Nonliving",
         "Other" = "allother",
         "Macroalgae Unsp" = "MacaUNSP", 
         "Seagrass" = "SEAGRASS",
         "Dictyota" = "DICT") %>%
  melt(id = ("Habitat")) %>%
  na.omit()
benthic$Habitat[benthic$Habitat == "Seagrass"] <- "Seagrass Fringe"


benthic_graph <-  ggplot(benthic, aes(x=Habitat, y= value, fill = variable))  +
  geom_col(position = "fill",
           width = (0.75)) +
  labs("Percent Cover", fill = "Biological Substrate") +
  xlab("Habitat Type") +
  ylab("Percent Cover") +
  theme(axis.title.y = element_text(size = 10)) 

benthic_graph

level_order_var <- c("Seagrass", "Other", "Sponge", "Coral", "Macroalgae Unsp", "Dictyota","Turf Algae","Non-living")
benthic$variable <- factor(benthic$variable, levels=level_order_var)
```


#Research Question 3: 
###Are there trends in habitat shifts?

```{r Ontogenetic Habitat Shifts}
#NCMRP DATASET
stjstt = getRvcData(years = 2017:2021, regions = c("STTSTJ"))
spcs = c("SPA CHRY")


den = getStratumDensity(stjstt, species = spcs, strata = c("AGRFSHLW","PTRFSHLW", "PVMTSHLW", "SCRSHLW", "HARDSHLW", "BDRKSHLW"), length_bins = seq(5,30,5)) %>%
  filter(length_class != "all")

den1 = getStratumDensity(stjstt, species = spcs, strata = c("AGRFSHLW","PTRFSHLW", "PVMTSHLW", "SCRSHLW", "HARDSHLW", "BDRKSHLW"), length_bins = seq(1,6,5)) %>%
  filter(length_class != "all")

filter0_5 <- den1 %>% filter(length_class == "[1, 6)")


den2 = getStratumDensity(stjstt, species = spcs, strata = c("AGRFSHLW","PTRFSHLW", "PVMTSHLW", "SCRSHLW", "HARDSHLW", "BDRKSHLW"), length_bins = seq(6,11,5)) %>%
  filter(length_class != "all")

filter6_10 <- den2 %>% filter(length_class == "[6, 11)")


den3 = getStratumDensity(stjstt, species = spcs, strata = c("AGRFSHLW","PTRFSHLW", "PVMTSHLW", "SCRSHLW", "HARDSHLW", "BDRKSHLW"), length_bins = seq(11,40,10)) %>%
  filter(length_class != "all")
filter11_40 <- den3 %>% filter(length_class != "< 11")

combined_df <- rbind(filter0_5, filter6_10, filter11_40)

updated <- combined_df %>%
  mutate(length_class = case_when(
    length_class == "[1, 6)" ~ "0-5", 
    length_class == "[6, 11)" ~ "6-10", 
    length_class == "[11, 21)" ~ "11-20",
    length_class == "[21, 31)" ~ "21-30",
    length_class == ">= 31" ~ ">= 31")
    )
hab <- updated %>%
  mutate(STRAT = case_when(
    STRAT == "AGRFSHLW" ~ "Coral Reef",
    STRAT == "PTRFSHLW" ~ "Coral Reef",
    STRAT == "SCRSHLW" ~ "Rocky Reef",
    STRAT == "BDRKSHLW" ~ "Rocky Reef",
    STRAT == "HARDSHLW" ~ "Coral Reef",
    STRAT == "PVMTSHLW" ~ "Hardbottom",
    TRUE ~ STRAT  # Keep other values as is
  ))

hab_fig_fill <- ggplot(hab, aes(x = length_class, y = density, fill = STRAT)) +
  geom_bar(position="fill", stat="identity", width = .75) +
  scale_x_discrete(limits = level_orderNCRMP) +
  labs(y=bquote("Density (N per 176 m"^"2"*")"),
       fill = "Habitat") +
  xlab("Size Class (cm)") +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("Hardbottom" = "#F8766D", "Rocky Reef" = "#00B0F6", "Coral Reef" = "purple")) +
  theme(axis.title.y = element_text(size = 9)) 
hab_fig_fill

level_orderNCRMP <- c("0-5", "6-10", "11-20", "21-30", ">= 31") 

hab$STRAT <- factor(hab$STRAT, levels = c("Hardbottom", "Rocky Reef", "Coral Reef"))



#TCRMP DATASET
tcrmp <- tcrmp %>%
  filter(Habitat != "Seagrass") %>%
  pivot_longer(cols = -Habitat, names_to = "size_class", values_to = "abundance")

level_orderTCRMP <- c("0_5", "6_10","11-20", "21_30", "31-40") 
tcrmp$Habitat<- factor(tcrmp$Habitat, levels = c("Hardbottom", "Rocky Reef", "Coral Reef"))

tcrmp_fig <- ggplot(tcrmp, aes(x = size_class, y = abundance, fill = Habitat)) +
  geom_bar(position="fill", stat="identity", width = .75) +
  scale_x_discrete(limits = level_orderTCRMP) +
  
  labs(y=bquote("Proportion of Total Density (N per 100 m"^"2"*")"),
       fill = "Habitat",
       title = "TCRMP Data + My Survey Data") +
  xlab("Size Class (cm)") +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("Hardbottom" = "#F8766D", "Rocky Reef" = "#00B0F6", "Coral Reef" = "purple")) +
  theme(legend.position = "top", 
        axis.title.y = element_text(size = 8)) 
tcrmp_fig

#PRELIMINARY DATASET
pre_long <- preliminarydata

level_order_prelim <- c("0-5", "6-10", "11-20", "21-30", "31-40") 
pre_long$Habitat<- factor(pre_long$Habitat, levels = c("hardbottom", "rocky reef", "coral reef"))

pre_fig <- ggplot(pre_long, aes(x = size_class, y = count, fill = Habitat)) +
  geom_bar(position="fill", stat="identity", width = .75) +
  scale_x_discrete(limits = level_order_prelim) +
  labs(y=bquote("Proportion of Total Abundance"),
       fill = "Habitat") +
  xlab("Size Class (cm)") +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("hardbottom" = "#F8766D", "rocky reef" = "#00B0F6", "coral reef" = "purple")) +
  theme(legend.position = "none", 
        axis.title.y = element_text(size = 9)) 
pre_fig

#COMBINED FIGURE
combined <- tcrmp_fig / hab_fig_fill / pre_fig
combined

```

### Research Question Extras
Kruskal Wallis on Depth: ###Is there a difference in depth between habitat types?
```{r Depth x Habitat Type}
depth <- nbr_complexdepth 


sum_depth<- ddply(depth, ~MajorHabitat, summarise, 
                      mean = mean(DepthMeters),
                       sd = sd(DepthMeters), 
                       n = length(DepthMeters), 
                       SEM = sd(DepthMeters)/sqrt(length(DepthMeters)))
view(sum_depth)
kruskal.test(DepthMeters ~ MajorHabitat, data = depth)
dunnTest(DepthMeters ~ MajorHabitat, data = depth,
         method="bonferroni")
pairwise.wilcox.test(depth$DepthMeters, depth$MajorHabitat,
                     p.adjust.method = "BH")


depth_fig <- ggplot(depth, aes(x = MajorHabitat, y = DepthMeters)) +
  geom_boxplot(outlier.colour = "grey", outlier.shape = 16, outlier.size = 1) +
  geom_jitter(width = 0.3, size = 1, alpha = 0.2) +
  labs(y=bquote("Depth (m)")) +
  xlab("Habitat Type") +
  theme(axis.text.x=element_text(size=9))
depth_fig
 
```

