---
title: "Data Management Final Project"
author: "By: Sierrah Mueller"
output: html_document
---

```{r setup libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(DHARMa)
library(MASS)
library(pscl)
library(AER)
library(tidyverse)
library(readxl)
library(sjPlot)
library(ggeffects)
library(reshape2)
library(plyr)
```

This document was created as a final project for Data Management/GIS. It highlights some of the statistical analyses conducted for Chapter 2 of my thesis titled: Filling Life History Gaps in a Data-Deficient Species, redtail parrotfish, _Sparisoma chrysopterum_, in St. Thomas, US Virgin Islands. Analysis broken down by research questions: 

 * __Research Question 1__ : Is there a difference in juvenile redtail parrotfish density between habitat types?
 * __Research Question 2__: How does the percent cover of benthic substrates across all habitat types affect the density of redtail parrotfish? 
 * __Research Question 3__: How do complexity and depth affect the density of redtail parrotfish and vary across habitat types surveyed?


*** 

## Research Question 1 {.tabset}
##### Is there a difference in juvenile redtail parrotfish density between habitat types?

### Data
```{r, echo = FALSE}
abundance_dfnewimport <- read_excel("../Data/CombinedData_Abundance-Habitat.xlsx", sheet = "OverallHabitatANOVA")
abundance_df <- abundance_dfnewimport %>%
  mutate(totalabun = rowSums(.[2:9]))
density_df <- abundance_df %>%
  mutate(density = totalabun*2) %>%
  mutate(MajorHabitat = ifelse(MajorHabitat == "Seagrass", "Seagrass Fringe", MajorHabitat))
head(density_df)
```

### Results
Kruskal Wallis test conducted to examine relationship between redtail parrotfish density by habitat type surveyed: hardbottom, rocky reef, and seagrass fringe. Kruskal Wallis test followed by a pairwise wilcox test for post-hoc analysis
```{r}

kruskal.test(density ~ MajorHabitat, data = density_df)
pairwise.wilcox.test(density_df$density, density_df$MajorHabitat,
                     p.adjust.method = "BH")

```

### Plot
```{r, echo= FALSE}
fig1 <- ggplot(density_df, aes(x = MajorHabitat, y = density)) +
  geom_boxplot(outlier.colour = "grey", outlier.shape = 16, outlier.size = 1) +
  geom_jitter(width = 0.3, size = 1, alpha = 0.2) +
  labs(y=bquote("Density (N per 100 m"^"2"*")")) +
  xlab("Habitat Type") +
  theme(axis.text.x=element_text(size=9))
fig1
```

Figure: Redtail parrotfish density (N per 100 m2) across hardbottom, rocky reef, and seagrass fringe habitats. 

*** 

## Research Question 2 {.tabset}
##### How does the percent cover of benthic substrates across all habitat types affect the density of redtail parrotfish? 

### Data
```{r, echo = FALSE }
glm_df <- read_excel("../Data/glm_df.xlsx")
head(glm_df)
```

### Results
Negative Binomial Regression (NBR) followed by likelihood ratio test to ensure NBR is a better fit than Poisson Model
```{r}
m_benthic <- glm.nb(Abundance ~ MacaUNSP + Dictyota + TURF + SEAGRASS + CORAL + SPONGE , data = glm_df)
summary(m_benthic)
odTest(m_benthic)
```


### Plot

```{r, echo= FALSE, message = FALSE, warning=FALSE}
#get confidence interval for complexity
CI_ben <- ggpredict(m_benthic, interval = "confidence", type = "fixed",
                 terms = "Dictyota")
CI_ben <- CI_ben %>% as.data.frame()

ggplot() +
  geom_point(data = glm_df, aes(x = Dictyota,
                               y = Abundance,
                              ),
             alpha = 0.2) +
  geom_path(data = CI_ben, aes(x = x, y = predicted)) +
  geom_ribbon(data = CI_ben, aes(x = x, ymin = conf.low,
                                  ymax = conf.high),
              alpha = 0.5) + 
  xlab("Percent Cover Dictyota spp.") +
  ylab("Predicted Abundance") +
  ggtitle("Relationship Between Percent Cover Dictyota & Juvenile Redtail Parrotfish Density")
```

Figure: Negative Binomial regression of the predicted density of redtail parrotfish in the presence of biological substrate Dictyota spp across all habitat types. 


***

## Research Question 3 {.tabset}
##### How do complexity and depth affect the density of redtail parrotfish and vary across habitat types surveyed?
```{r, include=FALSE}
nbr_complexdepth <- read_excel("../Data/nbr_complexdepth.xlsx")
noout <- subset(nbr_complexdepth,Complexity < 100)
complex_sum  <- ddply(nbr_complexdepth, ~MajorHabitat, summarise,
                      mean = mean(Complexity), 
                      sd = sd(Complexity), 
                      n = length(Complexity), 
                      SEM = sd(Complexity)/sqrt
                      (length(Complexity)))
```

### Data
```{r, echo = FALSE }
nbr_complexdepth <- read_excel("../Data/nbr_complexdepth.xlsx")
head(nbr_complexdepth)
```

### Results
#### __Part 1: Negative Binomial Regression (NBR)__
Examine the relationship between habitat characteristics complexity and depth on juvenile redtail parrotfish density
```{r NBR: ComplexityxDepth}
m_depthcomplex <- glm.nb(Density ~ Complexity + AvgDepth, data = noout)
summary(m_depthcomplex)
odTest(m_depthcomplex)

```

#### __Part 2: Kruskal Wallis Test__
Examine differences in topographic complexity between habitat types followed by post hoc analysis using pairwise wilcox test 
```{r}
kruskal.test(Complexity ~ MajorHabitat, data = nbr_complexdepth)
pairwise.wilcox.test(nbr_complexdepth$Complexity, nbr_complexdepth$MajorHabitat, 
                     p.adjust.method = "BH")

```


### Plots
```{r, echo= FALSE, message = FALSE, warning=FALSE}
CI_comp <- ggpredict(m_depthcomplex, interval = "confidence", type = "fixed",
                 terms = "Complexity")
CI_comp <- CI_comp %>% as.data.frame()

ggplot() +
  geom_point(data = noout, aes(x = Complexity,
                               y = Density,
                              ),
             alpha = 0.2) +
  geom_path(data = CI_comp, aes(x = x, y = predicted)) +
  geom_ribbon(data = CI_comp, aes(x = x, ymin = conf.low,
                                  ymax = conf.high),
              alpha = 0.5) +
  xlab("Topographic Complexity (cm)") +
  ylab("Predicted Abundance") +
  ggtitle("Relationship Between Topographic Complexity & Juvenile Redtail Parrotfish Density")
```


```{r, echo= FALSE}
fig_complex <- ggplot(nbr_complexdepth, aes(x = MajorHabitat, y = Complexity)) +
  geom_boxplot(outlier.colour = "grey", outlier.shape = 16, outlier.size = 1) +
  geom_jitter(width = 0.3, size = 1, alpha = 0.2) +
  labs(y=bquote("Topographic Complexity (cm)")) +
  xlab("Habitat Type") +
  theme(axis.text.x=element_text(size=9)) +
  ggtitle("Topographic Complexity Across Habitat Types")

fig_complex
```

